Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-06 WSSSPE1_report/WSSSPE1_summary_report.tex
more work is sustainability section - still in progress

Katz and Proctor~\cite{Katz_WSSSPE} included a set of questions that could
be used to measure software sustainability, and though these questions might falsely lead
to yes or no answers, the complete set would determine a range of values for
sustainability. These question are:
Will the software provide same functionality going forward even if environment changes?
Are the functionality and usability clear?
Will functionality be correct going into the future even if environment changes?
Is the functionality that which is desired by current and future users?
Can the software incorporate new things as they develop?
\todo{check paper for wording - might be better there than in this version, which comes from bullets}

\todo{See slide 9 of \url{http://dx.doi.org/10.6084/m9.figshare.853817} for summary table - incorporate/include table?}

>>>>

Katz and Proctor~\cite{Katz_WSSSPE} included a set of questions that could
be used to measure software sustainability, and though these questions might falsely lead
to yes or no answers, the complete set would determine a range of values for
sustainability. These question are:
\begin{itemize}
\item Will the software continue to provide the same functionality in the future, even if the environment (other parts of the infrastructure) changes?
\item Is the functionality and usability clearly explained to new users? Do users have a mechanism to ask questions and to learn about the element?
\item Will the functionality be correct in the future, even if the environment changes?
\item Does it provide the functionality that existing and future users want?
\item Does it incorporate new science/theory/tools as they develop?
\end{itemize}

Lenhardt~\cite{lenhardt-wssspe1-panel} summarized the contributions of the Modeling Sustainability papers as in Table~\ref{tab:defining-sustainability}. \todo{discussion of this?}

\begin{table}[h]
  \begin{scriptsize}
    \begin{center}
      \caption{Summary of Modeling Sustainability papers from Policy Panel.  Adapted from \cite{lenhardt-wssspe1-panel}}
      \label{tab:defining-sustainability}
      \begin{tabular}{|p{2.3cm}|p{3.6cm}|p{4.4cm}|p{4.8cm}|}
		\hline
{\bf Authors}
& {\bf Software}
& {\bf Sustainability}
& {\bf Approach to Understand or Evaluate Sustainability} \\
		\hline
Calero, et al. \cite{Calero_WSSSPE}
& General notion of software. Not explicitly defined.
& Sustainability is linked to quality.
& Add to ISO \\
		\hline
Venters, et al. \cite{Venters_WSSSPE}
& Software as science software; increasingly complex; service-oriented computing
& Extensibility, interoperability, maintainability, portability, reusability, scalability, efficiency
& Use various architecture evaluation approaches to assess sustainability \\
		\hline
Pierce, et al. \cite{Pierce_WSSSPE}
& Cyberinfrastructure software
& Sustainable to the extent to which there is a community to support it
& Open community governance \\
		\hline
Katz and Proctor \cite{Katz_WSSSPE}
& E-research infrastructures (i.e. cyberinfrastructure)
& Persisting over time, meeting original needs and projected needs
& Equates models for the creation of software with sustaining software \\
		\hline
Lenhardt, et al. \cite{Lenhardt_WSSSPE}
& Broadly defined as software supporting science
& Re-use; reproducible science
& Comparing data management life cycle to software development life cycle \\
		\hline
Weber, et al. \cite{Weber_WSSSPE}
& Software broadly defined; a software ecosystem
& Software niches
& Ecological analysis and ecosystem \\
		\hline
     \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}
